---
title: "EDS241: Assignment 4"
author: "Wylie Hampson"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 
  
``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}

# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", "here", "tidyr", "dplyr","stringr", "janitor", "here", "RDocumentation", "car",
           "cowplot", "ggplot2", "tinytex", "datasets", "tibble", "estimatr", "AER")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation


```

\noindent
This question will ask you to estimate the price elasticity of demand for fresh sardines across 56
ports located in 4 European countries with monthly data from 2013 to 2019. The data are
contained in the file EU_sardines.csv, which is available on Gauchospace.

\noindent
Each row in the data file is a combination of port location (where the fish is landed and sold) in a
given year and month. You can ignore the fact that the sample is not balanced (the number of
monthly observations varies across ports).

\noindent
For the assignment, you will need the following variables: year, month, country, port (port where
sardines are landed and sold), price_euro_kg (price per kg in €), and volume_sold_kg (quantity
of sardines sold in kg). In the questions below, I use log() to denote the natural logarithm.

\noindent
**Read in the data:**

```{r}
sardines <- read.csv(here("data", "EU_sardines.csv")) %>% 
  mutate(log_price_euro_kg = log(price_euro_kg), log_volume_sold_kg = log(volume_sold_kg))
```

\noindent
**Question a: Estimate a bivariate regression of log(volume_sold_kg) on log(price_euro_kg). What is the price elasticity of demand for sardines? Test the null hypothesis that the price elasticity is equal to -1.**

```{r}
model_a <- lm_robust(data = sardines, log_volume_sold_kg ~ log_price_euro_kg)

model_a_table <- broom::tidy(model_a) %>% 
  dplyr::select(term, estimate, std.error, p.value,) %>%
  knitr::kable()

model_a_table

linearHypothesis(model_a, c("log_price_euro_kg = -1"), white.adjust = "hc2")
```

\noindent
*The price elasticty of sardines is `r round(model_a$coefficients[2], 2)`. Because our linear hypothesis gives us a very low p-value under 0.001, we can reject the null hypthesis that the price eleasticity is equal to -1.*

\noindent
**Question b: Like in Lecture 8 (see the IV.R script), we will use wind_m_s as an instrument for log(price_euro_kg). To begin, estimate the first-stage regression relating log(price_euro_kg) to wind_m_s. Interpret the estimated coefficient on wind speed. Does it have the expected sign? Also test for the relevance of the instrument and whether it is a “weak” instrument by reporting the proper F-statistic. **

```{r}
model_b <- lm_robust(data = sardines, formula = log_price_euro_kg ~ wind_m_s)

model_b_table <- broom::tidy(model_b) %>% 
  dplyr::select(term, estimate, std.error, p.value) %>%
  knitr::kable()

model_b_table

summary(model_b)
```

\noindent
*The estimated coefficient for wind speed on log(price/kg) is `r round(model_b$coefficients[2], 2)`, meaning that for every 1 percent increase of wind speed, the log(price/kg) in euros will increase by an estimated `r round(model_b$coefficients[2], 2)` percentage points. This is the sign that we would expect to see because higher windspeed would make fishing more difficult, which would raise the price of fish. The F-statistic for wind speed is `r model_b$fstatistic[1]`, so it is a relevant and non-weak instrument.*

\noindent
**Question c: Estimate the TSLS estimator of the price elasticity of demand for sardines using wind_m_s as an instrument for log(price_euro_kg). What is the estimated price elasticity of demand for sardines?**

```{r}
tsls <- ivreg(log_volume_sold_kg ~ log_price_euro_kg | wind_m_s, data = sardines)
summary(tsls)

tsls_table <- broom::tidy(tsls) %>% 
  dplyr::select(term, estimate, std.error, p.value) %>%
  knitr::kable()

tsls_table
```

\noindent
*The 2 stage least square estimator for price elasticity of demarnd for sardines, using wind speed as an instrument for log(price_euro_kg), is `r tsls$coefficients[2]`. *

\noindent
**Question d:  Repeat the exercise in (c), but include fixed effects for each year, month, and country. [Hint: you can use the command “as.factor(country) + as.factor(year) +as.factor(month)” to the ivreg function in R]. Report the estimated price elasticity of demand and the F-statistic testing for relevant and non-weak instruments.**

```{r}
tsls_d <- ivreg(log_volume_sold_kg ~ log_price_euro_kg +
                as.factor(country) +
                as.factor(year) +
                as.factor(month) | wind_m_s +
                as.factor(country) +
                as.factor(year) +
                as.factor(month),
                data = sardines)

summary(tsls_d)

tsls_d_table <- broom::tidy(tsls_d) %>% 
  dplyr::select(term, estimate, std.error, p.value) %>%
  knitr::kable()

tsls_d_table

# To find the F-statistic
model_d <- lm_robust(log_price_euro_kg ~ wind_m_s +
                       as.factor(country) +
                       as.factor(year) +
                       as.factor(month), 
                     data = sardines)

model_d_test <- linearHypothesis(model_d, "wind_m_s = 0", white.adjust = "hc2")

model_d_test

```

\noindent
*Using fixed effects for each year, country, and month, the estimated price elasticity of demand is `r round(tsls_d$coefficient[2], 2)`. The F-statistic is `r round(model_d_test$Chisq[2], 2)`, meaning these are not a weak instruments.*

